---
title: "Debris analysis using soupX"
output:
  html_document: default
  pdf_document: default
---


```{r}
library(dplyr)
library(Seurat)
library(SC3)
library(SingleCellExperiment)
#install.packages("githubinstall")
#library(githubinstall)
#githubinstall("scmap")
library(scmap)
#devtools::install_github("constantAmateur/SoupX")
library(SoupX)
library(ggplot2)
library(tidyr)

```

The starting point is to download the raw and filtered cellranger output and extract them to a folder.
#Loading the data
SoupX comes with a convenience function for loading 10X data processed using cellranger.


#Build data table
Read in every 10x sample by soupX tool, split data with regards to reads in 0, 1, 2, 3-100 (debris) and 101-10000 (cellular) nUMI intervals, save nUMI information for 3-100 (debris) and 101-10000 (cellular) droplets.

```{r}
samples=read.csv("/home/jovyan/tisstab/analysis/allsamples.csv")
#Exclude bad sample
samples=samples[samples$sample!="HCATisStabAug177276393",]



nDrop.all=data.frame(nUMI=numeric(),Donor=character(), Time=character, sample.id=character(), Tissue=character())
nDrop.all.hi=data.frame(nUMI=numeric(),Donor=character(), Time=character, sample.id=character(), Tissue=character())
samples$prop0=0
samples$prop1=0
samples$prop2=0
samples$prop3_100=0
samples$prop100_higher=0

for (i in 1:nrow(samples)){
  dataDirs = c(paste0("/home/jovyan/tisstab/data/", samples$sample[i]))
  scl = load10X(dataDirs, keepDroplets = TRUE)
  scl$channels$Channel1 = estimateSoup(scl$channels$Channel1)
  nDrop=as.data.frame(scl$channels$Channel1$nDropUMIs[((scl$channels$Channel1$nDropUMIs <=100)& (scl$channels$Channel1$nDropUMIs>1))])
  colnames(nDrop)="nUMI"
  nDrop$Donor=samples$patient[i]
  nDrop$Time=samples$timepoint[i]
  nDrop$sample.id=samples$sample[i]
  nDrop$Tissue=samples$organ[i]
  nDrop.all=rbind(nDrop.all, nDrop)
  
  nDrop.hi=as.data.frame(scl$channels$Channel1$nDropUMIs[((scl$channels$Channel1$nDropUMIs >100)& (scl$channels$Channel1$nDropUMIs<=10000))])
  nDrop.hi$Donor=samples$patient[i]
  nDrop.hi$Time=samples$timepoint[i]
  nDrop.hi$sample.id=samples$sample[i]
  nDrop.hi$Tissue=samples$organ[i]
  nDrop.all.hi=rbind(nDrop.all.hi, nDrop.hi)
  
  leng=length(scl$channels$Channel1$nDropUMIs)
  samples$prop0[i]=length(scl$channels$Channel1$nDropUMIs[scl$channels$Channel1$nDropUMIs==0]) / leng
  samples$prop1[i]=length(scl$channels$Channel1$nDropUMIs[scl$channels$Channel1$nDropUMIs==1]) / leng
  samples$prop2[i]=length(scl$channels$Channel1$nDropUMIs[scl$channels$Channel1$nDropUMIs==2]) / leng
  samples$prop3_100[i]=length(scl$channels$Channel1$nDropUMIs[(scl$channels$Channel1$nDropUMIs>2) & (scl$channels$Channel1$nDropUMIs<=100) ]) / leng
  samples$prop100_higher[i]=length(scl$channels$Channel1$nDropUMIs[scl$channels$Channel1$nDropUMIs>100]) / leng
}

saveRDS(nDrop.all.hi, "/home/jovyan/tisstab/analysis/soupX.nDropUMI.hi.rds")
saveRDS(nDrop.all, "/home/jovyan/tisstab/analysis/soupX.nDropUMI.rds")
saveRDS(samples, "/home/jovyan/tisstab/analysis/soupX.samples.rds")


```

```{r data summary function}
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = mad(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}
```


#Observe proportions of droplets by intervals of nUMI

```{r}
samples=readRDS("/home/jovyan/tisstab/analysis/soupX.samples.rds")



#Percentages from full data

data_long <- gather(samples, Interval, Proportion, prop0:prop100_higher, factor_key=TRUE)
df3=data_summary(data_long, varname="Proportion", 
                    groupnames=c("Interval", "timepoint", "organ"))

ggplot(df3[df3$organ=="Spleen",], aes(x=timepoint, y=Proportion, group=Interval, color=Interval)) + 
   geom_errorbar(aes(ymin=Proportion-sd, ymax=Proportion+sd), width=0.2, position=position_dodge(0.3), size=1) +
   scale_color_grey()+
    geom_line(aes(lty=Interval),position=position_dodge(0.3), size=1.5) + geom_point(position=position_dodge(0.3), size=4)+theme_classic()+
  geom_point(data=df3[df3$organ=="Spleen",], aes(x=timepoint, y=Proportion, group=Interval),position=position_dodge(0.3) )

ggplot(df3[df3$organ=="Lung",], aes(x=timepoint, y=Proportion, group=Interval, color=Interval)) + 
   geom_errorbar(aes(ymin=Proportion-sd, ymax=Proportion+sd), width=0.2, position=position_dodge(0.3), size=1) +
   scale_color_grey()+
    geom_line(aes(lty=Interval),position=position_dodge(0.3), size=1.5) + geom_point(position=position_dodge(0.3), size=4)+theme_classic()+
  geom_point(data=df3[df3$organ=="Lung",], aes(x=timepoint, y=Proportion, group=Interval),position=position_dodge(0.3) )

ggplot(df3[df3$organ=="Oesophagus",], aes(x=timepoint, y=Proportion, group=Interval, color=Interval)) + 
   geom_errorbar(aes(ymin=Proportion-sd, ymax=Proportion+sd), width=0.2, position=position_dodge(0.3), size=1) +
   scale_color_grey()+
    geom_line(aes(lty=Interval),position=position_dodge(0.3), size=1.5) + geom_point(position=position_dodge(0.3), size=4)+theme_classic()+
  geom_point(data=df3[df3$organ=="Oesophagus",], aes(x=timepoint, y=Proportion, group=Interval),position=position_dodge(0.3) )



```

```{r}

timecolors=readRDS("/home/jovyan/tisstab/timecolors.rds" )
donorcolors=readRDS("/home/jovyan/tisstab/donorcolors.rds")

nDrop.all=readRDS("/home/jovyan/tisstab/analysis/soupX.nDropUMI.rds")


#Plot spleen only
nDrop=nDrop.all[nDrop.all$Tissue=="Spleen",]

ggplot(nDrop, aes(x=Donor, y=nUMI, fill=Time)) +
geom_boxplot(outlier.shape=".")+
  theme_classic()+  scale_fill_manual(values=timecolors)+ theme_classic()
  
ggplot(nDrop, aes(x=Time, y=nUMI, fill=Donor)) +
geom_boxplot(outlier.shape=".")+
  theme_classic()+    scale_fill_manual(values=donorcolors[1:5])+ theme_classic()



#All tissues debris
nDrop.all=nDrop.all[nDrop.all$nUMI>2,]

samples$mean=0
samples$median=0
samples$q75=0
samples$stdev=0
for (i in (1:nrow(samples))){
  mea=mean(nDrop.all$nUMI[nDrop.all$sample==samples$sample[i]])
  samples$mean[i]=mea
  quant=quantile(nDrop.all$nUMI[nDrop.all$sample==samples$sample[i]], c(.5, .75))
  samples$median[i]=quant[1]
  samples$q75[i]=quant[2]
  samples$stdev[i]=sd(nDrop.all$nUMI[nDrop.all$sample==samples$sample[i]])

}


df3 <- data_summary(samples, varname="mean", 
                    groupnames=c("organ", "timepoint"))

head(df3)
# Standard deviation of the mean plotted
ggplot(df3, aes(x=timepoint, y=mean, group=organ, color=organ)) + 
   geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.2, position=position_dodge(0.3), size=1) +
    geom_line(position=position_dodge(0.3), size=1.5) + geom_point(position=position_dodge(0.3), size=4)+theme_classic()+
  geom_point(data=samples, aes(x=timepoint, y=mean, group=organ),position=position_dodge(0.3) )


spleen=samples[samples$organ=="Spleen",]
#T-tests
#0h vs 72h
t.test(spleen$mean[spleen$timepoint=="72h"], y = spleen$mean[spleen$timepoint=="0h"], paired=TRUE, alternative = "g")
#0h vs 24h
t.test(spleen$mean[spleen$timepoint=="24h"], y = spleen$mean[spleen$timepoint=="0h"], paired=FALSE, alternative = "g")


```


#All tissues cellular material

```{r}

hisamples=samples

nDrop.all.hi=readRDS("/home/jovyan/tisstab/analysis/soupX.nDropUMI.hi.rds")
colnames(nDrop.all.hi)[1]="nUMI"
hisamples$mean=0
hisamples$median=0
hisamples$q75=0
hisamples$stdev=0

for (i in (1:nrow(hisamples))){
  mea=mean(nDrop.all.hi$nUMI[nDrop.all.hi$sample==hisamples$sample[i]])
  hisamples$mean[i]=mea
  quant=quantile(nDrop.all.hi$nUMI[nDrop.all.hi$sample==hisamples$sample[i]], c(.5, .75))
  hisamples$median[i]=quant[1]
  hisamples$q75[i]=quant[2]
  hisamples$stdev[i]=sd(nDrop.all.hi$nUMI[nDrop.all.hi$sample==hisamples$sample[i]])

}


df3 <- data_summary(hisamples, varname="mean", 
                    groupnames=c("organ", "timepoint"))

head(df3)
# Standard deviation of the mean plotted
ggplot(df3, aes(x=timepoint, y=mean, group=organ, color=organ)) + 
   geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=0.2, position=position_dodge(0.3), size=1) +
    geom_line(position=position_dodge(0.3), size=1.5) + geom_point(position=position_dodge(0.3), size=4)+theme_classic()+
  geom_point(data=hisamples, aes(x=timepoint, y=mean, group=organ),position=position_dodge(0.3) )


spleen=hisamples[hisamples$organ=="Spleen",]
#T-tests
#0h vs 72h
t.test(spleen$mean[spleen$timepoint=="72h"], y = spleen$mean[spleen$timepoint=="0h"], paired=TRUE, alternative = "l")
#0h vs 24h
t.test(spleen$mean[spleen$timepoint=="24h"], y = spleen$mean[spleen$timepoint=="0h"], paired=FALSE, alternative = "l")
```





```{r}
sessionInfo()
```
